The Development of Abstract Concepts in Children's Early Lexical Networks

### 2019-07-16 CMCL/ms.Rmd

<!--On the other hand, children can track co-occurrence statistics \cite{saffran1996} and there is evidence that they use co-occurrence to make category-based generalization \cite{fisher2011,matlen2015}. 
Here we investigate the way abstract concepts may develop from the interaction of the children developing lexicon and co-occurrence information in the caregiver's speech. 
-->
We study the real conceptual development (i.e. as induced by the real trajectory of word learning) in comparison to two hypothetical developmental scenarios induced by two divergent proposals about the trajectory of word learning. On the first proposal, past lexical knowledge facilitates the future learning of related words, e.g., the word "cat" is more likely to be followed by another animal name than it is to be followed by a food name \cite{steyvers2005,borovsky2016}. On the second proposal, past lexical knowledge does not influence future learning, e.g., learning the word "cat" does not necessarily increase the odds that the next word will be another animal name  \cite{hills2009,sizemore2018}.

The paper is organized as follows. First, we describe the research strategy. In brief, we represented the developing lexicon as an evolving network and we used word co-occurrence in parent speech as a measure of words' relatedness. We operationalized abstract concepts as the highly interconnected regions of the network.  Second, we explore how the pattern of children's word learning influences higher-level conceptual development, and whether this development corresponds to a simultaneous or a sequential conceptual growth.

@@

The networks' nodes were nouns from Wordbank \cite{frank2017}, an open repository aggregating cross-linguistic developmental data of the MacArthur-Bates Communicative Development Inventory (CDI), a parent report vocabulary checklist, Toddler version \cite{fenson94}. Pairs of nouns were linked by weighted edges representing their semantic similarity derived based on co-occurrence in the corpus of child-directed speech CHILDES \cite{macwhinney2014}, using Word2Vec algorithm \cite{mikolov2013}.

First, we constructed the end-state network based on all nouns learned by the last age of acquisition. We used a subset of CDI nouns for which cross-linguistic translations are present, allowing us to explore cross-linguistic variability. We used data from the ten following languages: Croatian, Danish, English, French, Italian, Norwegian, Russian, Spanish, Swedish, and Turkish. The size of this subset varied from `r max_n` in Russian (representing `r max_per`% of total nouns in this language) to `r min_n` in Turkish (representing `r min_per`% of total nouns). Second, in order to study development towards the end-state, we constructed a different network at each month, based on the nouns that have been learned by that month.

@@

$$
R(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fn(\mathcal{C}_t)|}
$$
Where $tp(\mathcal{C}_t)$ are the true positives, defined as the word pairs that are placed in the same cluster under $\mathcal{C}_t$ and in the same cluster under $\mathcal{C}^*$. $fp(\mathcal{C}_t)$ are the false positives defined as the pairs placed in the same cluster under $\mathcal{C}_t$ and in different clusters under $\mathcal{C}^*$. Finally, $fn(\mathcal{C}_t)$ are the false negatives defined as the pairs placed in different clusters under $\mathcal{C}_t$ and in the same cluster under $\mathcal{C}^*$.

>>>>

<!--On the other hand, children can track co-occurrence statistics \cite{saffran1996} and there is evidence that they use co-occurrence to make category-based generalization \cite{fisher2011,matlen2015}. 
Here we investigate the way abstract concepts may develop from the interaction of the children developing lexicon and co-occurrence information in the caregiver's speech. 
-->
We study the real conceptual development (i.e. as induced by the real trajectory of word learning) in comparison to two hypothetical developmental scenarios induced by two possible mechanisms of word learning. On the first mechanism, past lexical knowledge facilitates the future learning of related words, e.g., the word "cat" is more likely to be followed by another animal name than it is to be followed by a food name \cite{steyvers2005,borovsky2016}. On the second mechanism, past lexical knowledge does not influence future learning, e.g., learning the word "cat" does not necessarily increase the odds that the next word will be another animal name  \cite{hills2009,sizemore2018}.

The paper is organized as follows. First, we describe the research strategy. In brief, we represented the developing lexicon as an evolving network and we used word co-occurrence in parent speech as a measure of words' relatedness. We operationalized abstract concepts as the highly interconnected regions of the network.  Second, we explore how the pattern of children's word learning influences higher-level conceptual development, and whether this development corresponds to simultaneous or sequential conceptual growth.

@@

The networks' nodes were nouns from Wordbank \cite{frank2017}, an open repository aggregating cross-linguistic developmental data of the MacArthur-Bates Communicative Development Inventory (CDI), Toddler version \cite{fenson94}. Pairs of nouns were linked by weighted edges representing their semantic similarity derived based on co-occurrence in the corpus of child-directed speech CHILDES \cite{macwhinney2014}, using the Word2Vec algorithm \cite{mikolov2013}.

First, we constructed the end-state network based on all nouns learned by the last age of acquisition. We used a subset of CDI nouns for which cross-linguistic translations are present, allowing us to explore cross-linguistic variability. We used data from the following ten languages: Croatian, Danish, English, French, Italian, Norwegian, Russian, Spanish, Swedish, and Turkish. The size of this subset varied from `r max_n` in Russian (representing `r max_per`% of total nouns present in the CDI data of this language) to `r min_n` in Turkish (representing `r min_per`% of total nouns). Second, in order to study development towards the end-state, we constructed a different network at each month, based on the nouns that have been learned by that month.

@@

$$
R(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fn(\mathcal{C}_t)|}
$$
Where $tp(\mathcal{C}_t)$ are the true positives, defined as the word pairs that are placed in the same cluster under $\mathcal{C}_t$ and in the same cluster under $\mathcal{C}^*$. $fp(\mathcal{C}_t)$ are the false positives,
defined as the pairs placed in the same cluster under $\mathcal{C}_t$ and in different clusters under $\mathcal{C}^*$. Finally, $fn(\mathcal{C}_t)$ are the false negatives, defined as the pairs placed in different clusters under $\mathcal{C}_t$ and in the same cluster under $\mathcal{C}^*$.

### 2019-07-16 CMCL/ms.tex

We study the real conceptual development (i.e.~as induced by the real
trajectory of word learning) in comparison to two hypothetical
developmental scenarios induced by two divergent proposals about the
trajectory of word learning. On the first proposal, past lexical
knowledge facilitates the future learning of related words, e.g., the
word ``cat'' is more likely to be followed by another animal name than
it is to be followed by a food name \cite{steyvers2005,borovsky2016}. On
the second proposal, past lexical knowledge does not influence future
learning, e.g., learning the word ``cat'' does not necessarily increase
the odds that the next word will be another animal name
\cite{hills2009,sizemore2018}.

The paper is organized as follows. First, we describe the research
strategy. In brief, we represented the developing lexicon as an evolving
network and we used word co-occurrence in parent speech as a measure of
words' relatedness. We operationalized abstract concepts as the highly
interconnected regions of the network. Second, we explore how the
pattern of children's word learning influences higher-level conceptual
development, and whether this development corresponds to a simultaneous
or a sequential conceptual growth.

@@

The networks' nodes were nouns from Wordbank \cite{frank2017}, an open
repository aggregating cross-linguistic developmental data of the
MacArthur-Bates Communicative Development Inventory (CDI), a parent
report vocabulary checklist, Toddler version \cite{fenson94}. Pairs of
nouns were linked by weighted edges representing their semantic
similarity derived based on co-occurrence in the corpus of
child-directed speech CHILDES \cite{macwhinney2014}, using Word2Vec
algorithm \cite{mikolov2013}.

First, we constructed the end-state network based on all nouns learned
by the last age of acquisition. We used a subset of CDI nouns for which
cross-linguistic translations are present, allowing us to explore
cross-linguistic variability. We used data from the ten following
languages: Croatian, Danish, English, French, Italian, Norwegian,
Russian, Spanish, Swedish, and Turkish. The size of this subset varied
from 314 in Russian (representing 100\% of total nouns in this language)
to 176 in Turkish (representing 59.26\% of total nouns). Second, in
order to study development towards the end-state, we constructed a
different network at each month, based on the nouns that have been
learned by that month.

@@

\[
R(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fn(\mathcal{C}_t)|}
\] Where \(tp(\mathcal{C}_t)\) are the true positives, defined as the
word pairs that are placed in the same cluster under \(\mathcal{C}_t\)
and in the same cluster under \(\mathcal{C}^*\). \(fp(\mathcal{C}_t)\)
are the false positives defined as the pairs placed in the same cluster
under \(\mathcal{C}_t\) and in different clusters under
\(\mathcal{C}^*\). Finally, \(fn(\mathcal{C}_t)\) are the false
negatives defined as the pairs placed in different clusters under
\(\mathcal{C}_t\) and in the same cluster under \(\mathcal{C}^*\).

>>>>

We study the real conceptual development (i.e.~as induced by the real
trajectory of word learning) in comparison to two hypothetical
developmental scenarios induced by two possible mechanisms of word
learning. On the first mechanism, past lexical knowledge facilitates the
future learning of related words, e.g., the word ``cat'' is more likely
to be followed by another animal name than it is to be followed by a
food name \cite{steyvers2005,borovsky2016}. On the second mechanism,
past lexical knowledge does not influence future learning, e.g.,
learning the word ``cat'' does not necessarily increase the odds that
the next word will be another animal name \cite{hills2009,sizemore2018}.

The paper is organized as follows. First, we describe the research
strategy. In brief, we represented the developing lexicon as an evolving
network and we used word co-occurrence in parent speech as a measure of
words' relatedness. We operationalized abstract concepts as the highly
interconnected regions of the network. Second, we explore how the
pattern of children's word learning influences higher-level conceptual
development, and whether this development corresponds to simultaneous or
sequential conceptual growth.

@@

The networks' nodes were nouns from Wordbank \cite{frank2017}, an open
repository aggregating cross-linguistic developmental data of the
MacArthur-Bates Communicative Development Inventory (CDI), Toddler
version \cite{fenson94}. Pairs of nouns were linked by weighted edges
representing their semantic similarity derived based on co-occurrence in
the corpus of child-directed speech CHILDES \cite{macwhinney2014}, using
the Word2Vec algorithm \cite{mikolov2013}.

First, we constructed the end-state network based on all nouns learned
by the last age of acquisition. We used a subset of CDI nouns for which
cross-linguistic translations are present, allowing us to explore
cross-linguistic variability. We used data from the following ten
languages: Croatian, Danish, English, French, Italian, Norwegian,
Russian, Spanish, Swedish, and Turkish. The size of this subset varied
from 314 in Russian (representing 100\% of total nouns present in the
CDI data of this language) to 176 in Turkish (representing 59.26\% of
total nouns). Second, in order to study development towards the
end-state, we constructed a different network at each month, based on
the nouns that have been learned by that month.

@@

\[
R(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fn(\mathcal{C}_t)|}
\] Where \(tp(\mathcal{C}_t)\) are the true positives, defined as the
word pairs that are placed in the same cluster under \(\mathcal{C}_t\)
and in the same cluster under \(\mathcal{C}^*\). \(fp(\mathcal{C}_t)\)
are the false positives, defined as the pairs placed in the same cluster
under \(\mathcal{C}_t\) and in different clusters under
\(\mathcal{C}^*\). Finally, \(fn(\mathcal{C}_t)\) are the false
negatives, defined as the pairs placed in different clusters under
\(\mathcal{C}_t\) and in the same cluster under \(\mathcal{C}^*\).
